Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-11-01 files/Overview.tex
overview v2

\subsubsection{Chapter 1}
This dissertation is a literature study that investigates the validity of different linear models. The term validity in this dissertation means how well a model compares to a more realistic model for real world applications. The first chapter of the dissertation introduces the models. The main model is the Timoshenko beam theory. The other models are a two-dimensional elastic model, a three-dimensional elastic model and a Reissner-Mindlin plate model. The goal of this dissertation to validate the use of the Timoshenko beam model and the Reissner-Mindlin plate model in applications. These models are simplified one-dimensional beam and two-dimensional plate models. But more realistic models exists, such as a two-dimensional and three-dimensional beam model and a three-dimensional plate model. Using modal analysis (see chapter 2), it is shown that the solutions of the models can be represented as a linear combination of the eigenvalues and eigenfunctions. Using this idea, it is only required to compare the eigenvalues and eigenfunctions of the models to be able to compare the difference between the solutions. In chapter 1, the models of the dissertation is given. They are given in a dimensionless form, and also the variational forms are derived. Model problems are then defined by including boundary conditions. These model problems are the models used in the rest of the dissertation.

\subsubsection{Chapter 2}
In this section, the existence of a unique solutions for a general vibration problem is investigated. The main article that is discussed is \cite{VV02}. In this article, the authors define a general vibration problem. The authors state four assumptions that need to be satified. Under these four assumptions, it can be shown that for the general vibration problem there exists a linear operator A, which is an infinitesimal generator of a $C_0$-semigroup of contractions with a domain in the problem space. This operator allows the general vibration problem to be rewritten as a system of first order differential equations $\displaystyle x' = Ax + f$. This differential equations are rewritten into an initial value problem which can be proven using semi-group theory to have a unique solution. The authors then provide the necessary proofs to show that a solution of the initial value problem is also a solution of the general vibration problem. An example of the application of this theory is provided using a cantilever Timoshenko beam.

@@

The second part of the chapter considers work done in a textbook \cite{SF97}. The specific work discussed, covers the convergence of the eigenvalues and eigenfunctions of a one-dimensional vibration problem when applying the Finite Element Method. The authors consider a general eigenvalue problem. Then using the Rayleigh-quotient, from the Rayleigh-Ritz method, as well as an approximation theorem from \cite{ORXX}, the main result is proven. The specific work done in this section is updating the notation of the textbook, as well as expanding some results so that the results are easier to understand.

\subsubsection{Chapter 4}
This chapter is a focus on the main model theory of this disseration, the Timoshenko beam theory. The first section is a discussion of modal analysis applied to the Timoshenko beam theory, and specifically a discussion of the article \cite{VV06}. In this article, the authors present a method to calculate the exact eigenvalues and eigenfunctions of a Timoshenko beam. This method considers the boundary conditions, which differs from most other research. Starting with a general eigenvalue problem for a Timoshenko beam model, the authors derive a general solution for this problem. The authors then explain the method by hands of an example by applying the method to a cantilever beam model. The next sections of this chapter also then looks at examples of applying the method, first to a cantilever beam model, and then to a free-free beam model. The examples shows the use of the method, and is preperation for chapter 6 where these results are required for the comparison of the models. The next section is a short discussion on a suspended beam model, afterwards an interesting article \cite{SP06} is discussed. In this article, the authors investigate the validity of the Timoshenko beam theory, by comparing the eigenvalues (natural frequencies) for a physical beam, to a Timoshenko beam and a three-dimensional beam using Finite Element Analysis. The authors suspended the beam, and vribrated the beam to induce a free-free beam and measure the natural frequencies. These emperical results are then compared to the theoretical results. This result, together with the results in chapter 6, give a good picture to the validity of the Timoshenko beam theory. Finally the chapter finished by looking at an interesting model of a cantilever beam with a tip-body with elastic interfaces. These results combine modal analysis and looks at the behavior of the model. This section is an example of the usefullness of the Timoshenko beam theory. A similar model using a more realistic two- or three-dimensional model will be very complex.

\subsubsection{Chapter 5}
In this chapter, the Finite Element Method is applied to cantilever two-dimensional,and three-dimensional cantilever beams and a cantilever Reissner-Mindlin plate. The aim of this section is to obtain a numerical formula to calculate the eigenvalues and eigenfunctions of the models. The results of these formulas will be used in chapter 6. The Finite Element Method is not applied to the Timoshenko beam theory, as chapter 4 provides an alternative method. For all the models, the Finite Element Method is applied using bi-cubic or tri-cubic basis functions to improve the rate of convergence and reduce the processing required to obtain accurate results. Each section ends with a eigenvalue problem for the models that can be easily applied to a computer program to calculate the eigenvalues and eigenvectors for the models. These numerical formulas will be used to obtain the results in chapter 6.

>>>>

\subsubsection{Chapter 1}
This dissertation is a literature study to investigate the validity of different linear models for beams and plates. The term validity in this dissertation means how well a model compares to a more realistic model for real world applications. The first chapter of the dissertation introduces the models. The simplest model is the Timoshenko model. The other models are a two-dimensional elastic beam model, a three-dimensional elastic beam model and a Reissner-Mindlin plate model. The aim in this dissertation to validate the use of the Timoshenko beam model and the Reissner-Mindlin plate model in applications. These models are simplified one-dimensional beam and two-dimensional plate models. But more realistic models exists, such as a two-dimensional and three-dimensional beam model and a three-dimensional plate model. Using modal analysis (see Chapter 2), it is shown that the solutions of the models can be represented as a linear combination of the modal solutions. Using this idea, it is only required to compare the eigenvalues and eigenfunctions of the models to be able to compare the difference between the solutions. In Chapter 1, the models of the dissertation is given (they are given in a dimensionless form).  Model problems are then defined by including boundary conditions. These model problems are the models used in the rest of the dissertation. Also the variational forms are derived which enables us to see the similarity between the models.

\subsubsection{Chapter 2}
First, a variational form for a model problem of a cantilever Timoshenko beam given. This model problem in variational form is then extended to complete function spaces, and in turn a weak variational problem is obtained. This weak variational problem is used as an example to explain the main theory of the chapter, from the article \cite{VV02}. In this article, the authors present a general weak variational form. The weak variational form of all the model problems of this dissertation are special cases of this general weak variational problem. The assumptions and results can be formulated and applied to all the applications.

In this section, the existence of a unique solutions for a general vibration problem is investigated. The main article that is discussed is \cite{VV02}. In this article, the authors define a general vibration problem. The authors state four assumptions that need to be satified. Under these four assumptions, it can be shown that for the general vibration problem there exists a linear operator A, which is an infinitesimal generator of a $C_0$-semigroup of contractions with a domain in the problem space. This operator allows the general vibration problem to be rewritten as a system of first order differential equations $\displaystyle x' = Ax + f$. This differential equations are rewritten into an initial value problem which can be proven using semi-group theory to have a unique solution. The authors then provide the necessary proofs to show that a solution of the initial value problem is also a solution of the general vibration problem. An example of the application of this theory is provided using a cantilever Timoshenko beam.

@@

The second part of the chapter considers work done in a textbook \cite{SF97}. The specific work discussed, covers the convergence of the eigenvalues and eigenfunctions of a one-dimensional vibration problem when applying the Finite Element Method. The authors consider a general eigenvalue problem. Then using the Rayleigh-quotient, from the Rayleigh-Ritz method, as well as an approximation theorem from \cite{OR97}, the main result is proven. The specific work done in this section is updating the notation of the textbook, as well as expanding some results so that the results are easier to understand.

\subsubsection{Chapter 4}
This chapter is a focus on the main model theory of this disseration, the Timoshenko beam theory. The first section is a discussion of modal analysis applied to the Timoshenko beam theory, and specifically a discussion of the article \cite{VV06}. In this article, the authors present a method to calculate the exact eigenvalues and eigenfunctions of a Timoshenko beam. This method considers the boundary conditions, which differs from most other research. Starting with a general eigenvalue problem for a Timoshenko beam model, the authors derive a general solution for this problem. The authors then explain the method by hands of an example by applying the method to a cantilever beam model. The next sections of this chapter also then looks at examples of applying the method, first to a cantilever beam model, and then to a free-free beam model. The examples shows the use of the method, and is preperation for chapter 6 where these results are required for the comparison of the models. The next section is a short discussion on a suspended beam model, afterwards an important article \cite{SP06} is discussed. In this article, the authors investigate the validity of the Timoshenko beam theory, by comparing the eigenvalues (natural frequencies) for a physical beam, to a Timoshenko beam and a three-dimensional beam using Finite Element Analysis. The authors suspended the beam, and vribrated the beam to induce a free-free beam and measure the natural frequencies. These emperical results are then compared to the theoretical results. This result, together with the results in chapter 6, give a good picture to the validity of the Timoshenko beam theory. Finally the chapter finished by looking at an interesting model of a cantilever beam with a tip-body with elastic interfaces. These results combine modal analysis and look at the behavior of the model. This section is an example of the usefullness of the Timoshenko beam theory. A similar model using realistic two- or three-dimensional model will be very challenging.

\subsubsection{Chapter 5}
In this chapter, the Finite Element Method is applied to cantilever two-dimensional,and three-dimensional cantilever beams and a cantilever Reissner-Mindlin plate. The aim of this section is to obtain an algorithm to calculate the eigenvalues and eigenfunctions of the models. The results of these formulas will be used in chapter 6. The Finite Element Method is not applied to the Timoshenko beam theory, as chapter 4 provides an alternative method. For all the models, the Finite Element Method is applied using bi-cubic or tri-cubic basis functions to improve the rate of convergence and reduce the processing required to obtain accurate results. Each section ends with a eigenvalue problem for the models that can be easily applied to a computer program to calculate the eigenvalues and eigenvectors for the models. These algorithms will be used to obtain the results in chapter 6.
